---
title: Statistical analysis
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
geometry: margin=1in
---

```{r setup, include = FALSE}
#rm(list=ls())
setwd("C:/Users/meglarse/GitHub/eco-evo-stoich")

# Load packages and dependencies
#require(aggregate)
require(reshape); require(nlme); require(vegan)
require(igraph); require(bipartite)
require(plotrix);require(pander)
require(ggplot2); require(gridExtra)
require(reshape);require(nlme)
require(synchrony); require(astsa)
require(xtable)
# for latex tables
require(stargazer)

# load dependencies
source("./bin/plot-ts.R")

```

```{r global-opts, include = FALSE}

#knitr options

knitr::opts_chunk$set(fig.width=5, fig.height=5,
                      echo=FALSE, warning=FALSE, message=FALSE)

#ggplot options

```

```{r startup, include = FALSE}
# Initialize the caption list

#dir.create("supporting_files/output")
#dir.create("supporting_files/data")
#dir.create("diagnostics")

# make file for printing captions
#captext()

```

# Project Summary
**Author**
M.L. Larsen, *Indiana University, Bloomington, IN*

**Collaborators**

J.T. Lennon, *Indiana University, Bloomington, IN*

S.W. Wilhelm, *University of Tennessee - Knoxville, Knoxville TN*


**Project questions**

1. Does resource stoichiometry affect the growth rate of *Synechococcus*?
2. How does resource stoichiometry alter ecological dynamics? 
3. Does stoichiometry alter phenotypic (co)evolution in cyanobacteria and phage?

**Data collection**

Briefly, all data for this project was collected during a long term continuous culture experimental evolution study with *Synechococcus* and SRIM-8 cyanomyophage.

For a complete description of the materials and methods for this repository, see Larsen *et al.* 2016.


Funding for this project was provided in part by the National Science Foundation, Michigan State University BEACON Center for Evolution in Action, and Indiana University.

\newpage
\tableofcontents
\newpage

```{r convert-matrix}
# Read in the Breadth of Resistance Data
bor <- read.csv("./data/20150710_BORb.csv", header = F)

# prevent factors
dat <- apply(bor, 2, function(x) as.character(x))


##Clean up data for analysis
#keep columns of data that have less than 50% NA
dat1 <- dat[, colSums(is.na(dat)) < length(dat[8:nrow(dat),1]) * 0.50]

# need to remove columns that are all zeros. suggests that cells did not grow


dat2 <- dat1

#process matrix
#remove bacteria information 
res2 <- dat2[8:dim(dat2)[[1]],]

nms <- dat2[7,1:7]
nms <- t(lapply(nms,function(x) {paste("P.",gsub(" ",".",x),sep="")}))
colnames(res2)[1:7] <- nms

res3 <- as.data.frame(res2)
#head(res3)

m1 <- melt(res3,id=1:7)
m1$variable <- as.character(m1$variable)
#head(m1)

# now grab column variables
res2b <- t(dat2[1:7,7:dim(dat2)[[2]]])
nmsb <- dat2[1:7,7]
nmsb <- t(lapply(nmsb,function(x) {paste("B.",gsub(" ",".",x),sep="")}))
colnames(res2b) <- nmsb
vars <- rownames(res2b)
res2b <- cbind(res2b,variable=vars)

#combine covariates
res4 <- merge(m1,res2b,by=c("variable"),all.x=T)

# move value column
res4 <- cbind(res4[,-which(colnames(res4)=="value")],infectivity=res4$value)

# drop the now unnecessary variable column
res4 <- res4[,-1]

# ensure that infectivity is numeric
res4$infectivity <- as.numeric(as.character(res4$infectivity))

# clean up labels for factors, example:
#levels(res4$P.cID)

# note that there shouldn't be two different versions of N2, 
#but some of them are followed by an extra space.

# this turns P.cID into character data, 
# finds space characters and removes them, overwriting P.cID in res4
res4$P.cID <- gsub(" ","",as.character(res4$P.cID))

# also happens in B.Time.point
res4$B.tm.pt <- gsub(" ","",as.character(res4$B.tm.pt))

# create matrix of infectivity proportions
# by looking at a specific nutrient treatment, for example.
#head(res4)

#This file includes control data too
#write.csv(res4, file = "./output/20150710_bor-processed.csv", row.names = F)
```

## Interaction Strengths
### Data Reduction
```{r dof-function}
# Use processed datafile for remaining analyses
res4 <- read.csv("./output/20150710_bor-processed.csv", header = T)

# make function to calculate TEMPORAL infectivity proportions
infect.prop<-function(x){
  x <- x[!is.na(x)] # remove NA's
  length(x[x == 0])/length(x)
	}
# infect.prop(c(0,0,1))
# infect.prop(c(0,NA,1))
```

```{r, eval = FALSE}
# separate the infection data so that each time point has a past, contemp, and future metric
# in each coevolving chemostat

#use longform bor matrix

#Remove the ancestral phage data 
res4 <- subset(res4,P.cID !="RIM8-1")
res4 <- subset(res4,B.cID !="WH7803")
res4 <- subset(res4,B.cID !="WH8101")


dat = res4[res4$B.trt == "T",]

    cid.list <- c()
    time.list <- c()
    past.list <- c()
    cont.list <- c()
    fut.list <- c()

for(i in unique(dat$B.cID)){
  for(j in unique(dat$B.tm.pt)){
    tmp = dat[dat$B.cID == i & dat$B.daynumber == j,]
    tmp = na.omit(tmp)
    past.dat = subset(tmp, tmp$P.daynumber < tmp$B.daynumber)
    p = length(past.dat[past.dat$infectivity == 1,])/length(past.dat$infectivity)
    cont.dat = tmp[tmp$P.daynumber == tmp$B.daynumber,]
    c = length(cont.dat[cont.dat$infectivity == 1,])/length(cont.dat$infectivity)
    future.dat = tmp[tmp$P.daynumber > tmp$B.daynumber,]
    f = length(future.dat[future.dat$infectivity == 1,])/length(future.dat$infectivity)
    
    cid.list <- append(cid.list, i)
    time.list <- append(time.list, j)
    past.list <- append(past.list, p)
    cont.list <- append(cont.list, c)
    fut.list <- append(fut.list, f)
    
    #if(cid == NL.list){
    #  l == "N"
    #} else {
    #  l == "P"
    #}
    #lim.list <- append(lim, l)
    
    #if(id == phage.list){
    #  t = "T"
    #}else{
    #  t = "C"
    #}
    #trt.list <- append(trt,t)
    
  }
}

dat2 <- data.frame(cid.list,time.list,past.list,cont.list,fut.list)

m1 <- melt(dat2, id = c(trt,lim,time))
c2 <- cast(m1)
p = ggplot(c2, aes(x = variable, y = value)) +
  geom_point(color = c("white","grey50","black"),alpha = 0.7) +
  geom_boxplot(color = c("white","grey50","black"), alpha = 0.7) +
  facet_grid(trt ~ lim)
print(p)
```

```{r NL_controls_dof}
#Initialize data storage
lim.list <- c(); cid.list <- c(); phage.time <- c(); bac.time <- c(); inf.prob <- c();b.cid.list <- c()

r = 1
l = 0

#Remove the ancestral phage data 
res4 <- subset(res4,P.cID !="RIM8-1")
res4 <- subset(res4,B.cID !="WH7803")
res4 <- subset(res4,B.cID !="WH8101")

# Subsetting by phage chemostat ID
cID = c("N2","N3","N5")
  
for(i in cID){
  r <- r + l
  res4b <- subset(res4,(P.cID == i | is.na(P.cID)) & (B.lim == "N" & B.trt == "C" | is.na(B.cID)))
  
  # pull out the relevant columns after subsetting as needed.
  res5 <- res4b[,c("P.daynumber","B.daynumber","infectivity")]
  
  # prevent factors; coerce to numeric
  res5 <- as.data.frame(apply(res5,2,function(x) as.numeric(as.character(x))))

  m2 <- melt(res5,id=c("P.daynumber","B.daynumber"))
  c2 <- cast(m2,P.daynumber~B.daynumber,fun.aggregate = infect.prop)
  c3 <- c2
  
  
  rownames(c2) <- c2$P.daynumber
  c2 <- c2[,-1]
          
  write.csv(c2, file = paste("./output/20150710_infmatrix",i,"Ncontrol",".csv",sep = ""), row.names = T)
  
  m3 <- melt(c3,id=c("P.daynumber","B.daynumber"))
  l <- r + length(m3[,1])
  
  cid.list <- append(cid.list,rep(as.character(i),length(m3[,1])))
  b.cid.list <- append(b.cid.list,rep(as.character(unique(res4b$B.cID)),length(m3[,1])))
  phage.time <- append(phage.time,m3$P.daynumber)
  bac.time <- append(bac.time,m3$B.daynumber)
  inf.prob <- append(inf.prob, m3$value)
  
  }


bor.cstat.Ncontrol <- data.frame(b.cid.list,cid.list,phage.time,bac.time,inf.prob)
```

```{r NL_coevo_dof}
#Initialize data storage
lim.list <- c(); cid.list <- c(); phage.time <- c(); bac.time <- c(); inf.prob <- c();b.cid.list <- c()

r = 1
l = 0

#Remove the ancestral phage data 
res4 <- subset(res4,P.cID !="RIM8-1")
res4 <- subset(res4,B.cID !="WH7803")
res4 <- subset(res4,B.cID !="WH8101")

cID = c("N2","N3","N5")
  
for(i in cID){
  r <- r + l
  res4b <- subset(res4,(P.cID == i | is.na(P.cID)) & (B.lim == "N" & B.trt == "T" | is.na(B.cID)))
  
  # pull out the relevant columns after subsetting as needed.
  res5 <- res4b[,c("P.daynumber","B.daynumber","infectivity")]
  
  # prevent factors; coerce to numeric
  res5 <- as.data.frame(apply(res5,2,function(x) as.numeric(as.character(x))))

  m2 <- melt(res5,id=c("P.daynumber","B.daynumber"))
  c2 <- cast(m2,P.daynumber~B.daynumber,fun.aggregate = infect.prop)
  c3 <- c2
  
  
  rownames(c2) <- c2$P.daynumber
  c2 <- c2[,-1]
          
  write.csv(c2, file = paste("./output/20150710_infmatrix",i,"Ncoevo",".csv",sep = ""), row.names = T)
  m3 <- melt(c3,id=c("P.daynumber","B.daynumber"))
  l <- r + length(m3[,1])
  
  cid.list <- append(cid.list,rep(as.character(i),length(m3[,1])))
  b.cid.list <- append(b.cid.list,rep(as.character(unique(res4b$B.cID)),length(m3[,1])))
  phage.time <- append(phage.time,m3$P.daynumber)
  bac.time <- append(bac.time,m3$B.daynumber)
  inf.prob <- append(inf.prob, m3$value)
  
  }

bor.cstat.Ncoevo <- data.frame(b.cid.list, cid.list,phage.time,bac.time,inf.prob)

```

```{r PL_control_dof}
#Initialize data storage
lim.list <- c(); cid.list <- c(); phage.time <- c(); bac.time <- c(); inf.prob <- c();b.cid.list <- c()

r = 1
l = 0

#Remove the ancestral phage data 
res4 <- subset(res4,P.cID !="RIM8-1")
res4 <- subset(res4,B.cID !="WH7803")
res4 <- subset(res4,B.cID !="WH8101")

cID = c("P2","P4","P5")
  
for(i in cID){
  r <- r + l
  res4b <- subset(res4,(P.cID == i | is.na(P.cID)) & (B.lim == "P" & B.trt == "C" | is.na(B.cID)))
  
  # pull out the relevant columns after subsetting as needed.
  res5 <- res4b[,c("P.daynumber","B.daynumber","infectivity")]
  
  # prevent factors; coerce to numeric
  res5 <- as.data.frame(apply(res5,2,function(x) as.numeric(as.character(x))))

  m2 <- melt(res5,id=c("P.daynumber","B.daynumber"))
  c2 <- cast(m2,P.daynumber~B.daynumber,fun.aggregate = infect.prop)
  c3 <- c2
  
  
  rownames(c2) <- c2$P.daynumber
  c2 <- c2[,-1]
          
  write.csv(c2, file = paste("./output/20150710_infmatrix",i,"Pcontrol",".csv",sep = ""), row.names = T)
  
  m3 <- melt(c3,id=c("P.daynumber","B.daynumber"))
  l <- r + length(m3[,1])
  
  cid.list <- append(cid.list,rep(as.character(i),length(m3[,1])))
  b.cid.list <- append(b.cid.list,rep(as.character(unique(res4b$B.cID)),length(m3[,1])))
  phage.time <- append(phage.time,m3$P.daynumber)
  bac.time <- append(bac.time,m3$B.daynumber)
  inf.prob <- append(inf.prob, m3$value)
  
  }


bor.cstat.Pcontrol <- data.frame(b.cid.list, cid.list,phage.time,bac.time,inf.prob)
```

```{r PL_coevo_dof}
#Initialize data storage
lim.list <- c(); cid.list <- c(); phage.time <- c(); bac.time <- c(); inf.prob <- c();b.cid.list <- c()

r = 1
l = 0

#Remove the ancestral phage data 
res4 <- subset(res4,P.cID !="RIM8-1")
res4 <- subset(res4,B.cID !="WH7803")
res4 <- subset(res4,B.cID !="WH8101")

cID = c("P2","P4","P5")
  
for(i in cID){
  r <- r + l
  res4b <- subset(res4,(P.cID == i | is.na(P.cID)) & (B.lim == "P" & B.trt == "T" | is.na(B.cID)))
  
  # pull out the relevant columns after subsetting as needed.
  res5 <- res4b[,c("P.daynumber","B.daynumber","infectivity")]
  
  # prevent factors; coerce to numeric
  res5 <- as.data.frame(apply(res5,2,function(x) as.numeric(as.character(x))))

  m2 <- melt(res5,id=c("P.daynumber","B.daynumber"))
  c2 <- cast(m2,P.daynumber~B.daynumber,fun.aggregate = infect.prop)
  c3 <- c2
  
  
  rownames(c2) <- c2$P.daynumber
  c2 <- c2[,-1]
          
  write.csv(c2, file = paste("./output/20150710_infmatrix",i,"Pcoevo",".csv",sep = ""), row.names = T)
  m3 <- melt(c3,id=c("P.daynumber","B.daynumber"))
  l <- r + length(m3[,1])
  
  cid.list <- append(cid.list,rep(as.character(i),length(m3[,1])))
  b.cid.list <- append(b.cid.list,rep(as.character(unique(res4b$B.cID)),length(m3[,1])))
  phage.time <- append(phage.time,m3$P.daynumber)
  bac.time <- append(bac.time,m3$B.daynumber)
  inf.prob <- append(inf.prob, m3$value)
  
  }

bor.cstat.Pcoevo <- data.frame(b.cid.list, cid.list,phage.time,bac.time,inf.prob)

```

```{r merge-dof-data}
#bor.cstat.Ncontrol
bor.cstat.Ncontrol$trt <- rep("C", length(bor.cstat.Ncontrol[,1]))
bor.cstat.Ncontrol$lim <- rep("N", length(bor.cstat.Ncontrol[,1]))

#bor.cstat.Ncoevo
bor.cstat.Ncoevo$trt  <- rep("T", length(bor.cstat.Ncoevo[,1]))
bor.cstat.Ncoevo$lim <- rep("N", length(bor.cstat.Ncoevo[,1]))

#bor.cstat.Pcontrol
bor.cstat.Pcontrol$trt  <- rep("C", length(bor.cstat.Pcontrol[,1]))
bor.cstat.Pcontrol$lim <- rep("P", length(bor.cstat.Pcontrol[,1]))

#bor.cstat.Pcoevo
bor.cstat.Pcoevo$trt <- rep("T", length(bor.cstat.Pcoevo[,1]))
bor.cstat.Pcoevo$lim <- rep("P", length(bor.cstat.Pcoevo[,1]))

bor.cstat <- rbind(bor.cstat.Ncontrol, bor.cstat.Ncoevo, bor.cstat.Pcontrol, bor.cstat.Pcoevo)
colnames(bor.cstat) <- c("BcID","cID","phage.time","bac.time","inf.prob","trt","lim")

```

```{r, include = FALSE, eval=FALSE}
#res4 <- read.csv("./output/20150707_bor-processed.csv", header = T)

#modified the interaction numbers to properly reflect temporal infection dynamics
res4 <- read.csv("./output/20150710_bor-processed.csv", header = T)

head(res4)

res4$P.trt <- c("T")

#Remove the ancestral phage data 
res4 <- subset(res4,P.cID !="RIM8-1")
res4 <- subset(res4,B.cID !="WH7803")
res4 <- subset(res4,B.cID !="WH8101")

```

```{r, include = FALSE, eval=FALSE}
#Initialize data storage for the calculation of temporal infection dynamics on controls
lim.list <- c(); cid.list <- c(); phage.time <- c(); bac.time <- c(); inf.prob <- c()

r = 1
l = 0

lim = unique(res4$P.lim)

for(i in lim){
  r <- r + l
  res4b <- subset(res4,(P.lim == i | is.na(P.cID)) & (B.lim == i & B.trt == "C" | is.na(B.cID)))
  
  # pull out the relevant columns after subsetting as needed.
  res5 <- res4b[,c("P.daynumber","B.daynumber","infectivity")]
  
  # prevent factors; coerce to numeric
  res5 <- as.data.frame(apply(res5,2,function(x) as.numeric(as.character(x))))

  m2 <- melt(res5,id=c("P.daynumber","B.daynumber"))
  c2 <- cast(m2,P.daynumber~B.daynumber,fun.aggregate = infect.prop)
  c3 <- c2
  
  
  rownames(c2) <- c2$P.daynumber
  c2 <- c2[,-1]
           
  #save the infection matrix for plotting
  write.csv(c2, file = paste("./output/20150710-infmatrix",i,"controls.csv",sep = ""), row.names = T)
  
  m3 <- melt(c3,id=c("P.daynumber","B.daynumber"))
  l <- r + length(m3[,1])
  
  cid.list <- append(cid.list,rep(as.character(i),length(m3[,1])))
  phage.time <- append(phage.time,m3$P.daynumber)
  bac.time <- append(bac.time,m3$B.daynumber)
  inf.prob <- append(inf.prob, m3$value)
  }

bor.cstat.controls <- data.frame(cid.list,phage.time,bac.time,inf.prob)
bor.cstat.controls$B.trt <- c("C")


#Initialize data storage for the calculation of temporal infection dynamics on coevolved hosts
lim.list <- c(); cid.list <- c(); phage.time <- c(); bac.time <- c(); inf.prob <- c()

r = 1
l = 0

lim = unique(res4$P.lim)

for(i in lim){
  r <- r + l
  res4b <- subset(res4,(P.lim == i | is.na(P.cID)) & (B.lim == i & B.trt == "T" | is.na(B.cID)))
  
  # pull out the relevant columns after subsetting as needed.
  res5 <- res4b[,c("P.daynumber","B.daynumber","infectivity")]
  
  # prevent factors; coerce to numeric
  res5 <- as.data.frame(apply(res5,2,function(x) as.numeric(as.character(x))))

  m2 <- melt(res5,id=c("P.daynumber","B.daynumber"))
  c2 <- cast(m2,P.daynumber~B.daynumber,fun.aggregate = infect.prop)
  c3 <- c2
  
  
  rownames(c2) <- c2$P.daynumber
  c2 <- c2[,-1]
          
  write.csv(c2, file = paste("./output/20150710-infmatrix",i,"coevo.csv",sep = ""), row.names = T)  
  
  m3 <- melt(c3,id=c("P.daynumber","B.daynumber"))
  l <- r + length(m3[,1])
  
  cid.list <- append(cid.list,rep(as.character(i),length(m3[,1])))
  phage.time <- append(phage.time,m3$P.daynumber)
  bac.time <- append(bac.time,m3$B.daynumber)
  inf.prob <- append(inf.prob, m3$value)
  }

#create a list for each type of bacterial isolate
bor.cstat.trt <- data.frame(cid.list,phage.time,bac.time,inf.prob)
bor.cstat.trt$B.trt <- c("T")


#Merge data frames
bor.cstat.trt <- rbind(bor.cstat.trt,bor.cstat.controls)

#Change column names
colnames(bor.cstat.trt) <- c("lim","phage.time","bac.time","inf.prob","trt")

#save data file       
#write.csv(c2, file = paste("./output/infmatrix",i,".csv",sep = ""), row.names = T)
  
#head(bor.cstat)
```

\newpage

### RMANOVA for Interaction Strengths
```{r dof-stats}
#str(bor.cstat)
bor.stat <- bor.cstat
#bor.stat$phage.time <- as.factor(bor.stat$phage.time)
#bor.stat$bac.time <- as.factor(bor.stat$bac.time)

bor.stat$trt <- as.factor(bor.stat$trt)
bor.stat$lim <- as.factor(bor.stat$lim)

```

```{r}
#Control vs treatment
model.ar <- lme(inf.prob ~ lim * phage.time * bac.time, random = ~1 | cID,
                 correlation = corAR1(form = ~1 | cID),
                 data = bor.stat[bor.stat$trt == "C",])

model.arma1 <- lme(inf.prob ~ lim * phage.time * bac.time, random = ~1 | cID,
                 correlation = corARMA(form = ~1 | cID, p = 1, q = 1),
                 data = bor.stat[bor.stat$trt == "C",])

model.arma2 <- lme(inf.prob ~ lim * phage.time * bac.time, random = ~1 | cID,
                 correlation = corARMA(form = ~1 | cID, p = 1, q = 2),
                 data = bor.stat[bor.stat$trt == "C",])
mod.comp = anova(model.ar, model.arma1, model.arma2)


model.ar <- lme(inf.prob ~ lim * phage.time * bac.time, random = ~1 | cID,
                 correlation = corAR1(form = ~1 | cID),
                 data = bor.stat[bor.stat$trt == "T",])

model.arma1 <- lme(inf.prob ~ lim * phage.time * bac.time, random = ~1 | cID,
                 correlation = corARMA(form = ~1 | cID, p = 1, q = 1),
                 data = bor.stat[bor.stat$trt == "T",])

model.arma2 <- lme(inf.prob ~ lim * phage.time * bac.time, random = ~1 | cID,
                 correlation = corARMA(form = ~1 | cID, p = 1, q = 2),
                 data = bor.stat[bor.stat$trt == "T",])
mod.comp = anova(model.ar, model.arma1, model.arma2)
anova(model.ar)

```

```{r dof-stats1}
model.ar <- lme(inf.prob ~ trt* lim * phage.time * bac.time, random = ~1 | cID,
                 correlation = corAR1(form = ~1 | cID),
                 data = bor.stat)

model.arma1 <- lme(inf.prob ~ trt* lim * phage.time * bac.time, random = ~1 | cID,
                 correlation = corARMA(form = ~1 | cID, p = 1, q = 1),
                 data = bor.stat)

model.arma2 <- lme(inf.prob ~ trt* lim * phage.time * bac.time, random = ~1 | cID,
                 correlation = corARMA(form = ~1 | cID, p = 1, q = 2),
                 data = bor.stat)
mod.comp = anova(model.ar, model.arma1, model.arma2)

sum.mod.arma2=summary(model.arma2)
#anova(model.arma2)

stats = anova.lme(model.arma2,type = "sequential", adjustSigma = F)
#write.csv(file = "./supporting-files/chpt-tables/SOI-RMANOVA.csv", stats, row.names = FALSE)
pander(mod.comp[,c(-1,-2,-3,-7)])

#captext = paste0("*Model covariate comparison for RMANOVA.*")
emphasize.strong.rows(which(sum.mod.arma2$tTable[,5] < 0.05, arr.ind = TRUE))
pander(sum.mod.arma2$tTable[,c(2,4,5)], justify = "left")

```

```{r, include = FALSE, eval = FALSE}
#Using data that includes the control hosts

#double check this information... need to have phage infectivity from each chemostat against every bacterium...
#This file is created in section 2.2.2
#bor.cstat = "./output/"

bor.stat <- bor.cstat.trt
bor.stat$phage.time <- as.factor(bor.stat$phage.time)
bor.stat$bac.time <- as.factor(bor.stat$bac.time)
bor.stat$trt <- as.factor(bor.stat$trt)
bor.stat$lim <- as.factor(bor.stat$lim)

#Run RMANOVA with different models
model.ar <- lme(inf.prob ~ trt * lim* phage.time * bac.time, random = ~1,
                 correlation = corAR1(form = ~1 ),
                 data = bor.stat)

model.arma <- lme(inf.prob ~ trt* lim * phage.time * bac.time,random = ~1,
                 correlation = corARMA(form = ~1, p = 1, q = 1),
                 data = bor.stat)

model.arma1 <- lme(inf.prob ~ trt* lim * phage.time * bac.time, random = ~1,
                 correlation = corARMA(form = ~1, p = 1, q = 2),
                 data = bor.stat)

#Select which model fits the data better
anova(model.ar,model.arma, model.arma1)

#summary(model.arma1)
#anova(model.arma1)
set.caption("RMANOVA table for infection data")
pander(model.arma1)
#phage.time*lim; p = 0.001
#phage.time*bac.time; p = 0.0029
#lim*phage.time*bac.time; p = 0.0085
```

### Network plots
```{r con+coevo}
Ncon <- read.csv("./output/20150710-infmatrixNcontrols.csv", header = F)
Ncon[1,1] <- 999
Ncon[,1] <- as.numeric(as.character(Ncon[,1]))
Ncon <- as.matrix(Ncon)

Ntrt <- read.csv("./output/20150710-infmatrixNcoevo.csv", header = F)
Ntrt[1,1] <- 999
Ntrt[,1] <- as.numeric(as.character(Ntrt[,1]))
Ntrt <- as.matrix(Ntrt)

Pcon <- read.csv("./output/20150710-infmatrixPcontrols.csv", header = F)
Pcon[1,1] <- 999
Pcon[,1] <- as.numeric(as.character(Pcon[,1]))
Pcon <- as.matrix(Pcon)

Ptrt <- read.csv("./output/20150710-infmatrixPcoevo.csv", header = F)
Ptrt[1,1] <- 999
Ptrt[,1] <- as.numeric(as.character(Ptrt[,1]))
Ptrt <- as.matrix(Ptrt)

#Ncon;Ntrt
#Pcon;Ptrt

#Add column to Ncon for 166; data does not exist for this time point
V8 <- c(166,NA,NA,NA,NA,NA)
Ncon <- cbind(Ncon,V8)


# save timings
phage.time <- as.numeric(c(Ncon[2:nrow(Ncon),1]))
#phage.time <- as.numeric(c(Ncon[2:nrow(Ncon),1],Ntrt[2:nrow(Ntrt),1]))
bacteria.time <- as.numeric(c(Ncon[1,2:ncol(Ncon)],Ntrt[1,2:ncol(Ntrt)]))

#Trim data to remove column and row labels
Ncon <- Ncon[2:dim(Ncon)[[1]], 2:(dim(Ncon))[[2]]]
Pcon <- Pcon[2:dim(Pcon)[[1]], 2:(dim(Pcon))[[2]]]
Ntrt <- Ntrt[2:dim(Ntrt)[[1]], 2:(dim(Ntrt))[[2]]]
Ptrt <- Ptrt[2:dim(Ptrt)[[1]], 2:(dim(Ptrt))[[2]]]


```

```{r coevo-pub/fig, fig.height=3.75, fig.width=6}


#png(filename = "./supporting-files/figures/fig3-CoevoDynamics.png",
#    width = 1140, height = 570, units = "px", pointsize = 17,
#    bg = "white", family = "sans", type = "cairo")

#pdf(file = "./supporting-files/chpt-figs/fig-pub_CoevoDynamics.pdf")

par(mfrow = c(1,2),mar = c(5,6,5,1) + 0.1)

#Ncon
#Ntrt

#Option 1
#Merge data from each matrix
#data.in <- rbind(Ncon, Ntrt)
#data.in <- as.matrix(data.in)

#generate a square adjacency matrix, creating a place to store the data
#mat <- matrix(0,nrow=sum(dim(data.in)),ncol=sum(dim(data.in)))
  
#nrows <- dim(data.in)[[1]]
#big.nrows <- dim(mat)[[2]]
  
# create the phage-bacteria edges between time points
#mat[1:nrows,I(nrows+1):big.nrows] <- data.in

# convert adjacency matrix into graph object
#g <- graph.adjacency(mat,mode = "directed",weighted = T)

# extract weights
#wts <- g2[[9]][[4]]$weight
#wts <- E(g)$weight

#Option 2
# construct and populate adjacency matrix
size <- 1*dim(Ncon)[[1]]+2*dim(Ntrt)[[2]]
mat <- matrix(0,nrow = size,ncol = size)
#mat

# for multiple bacteria, single phage 
mat[1:dim(Ncon)[[1]],I(dim(Ncon)[[1]]+1):I(dim(Ncon)[[1]]+dim(Ncon)[[2]])] <-as.matrix(Ntrt)
mat[1:dim(Ncon)[[1]],I(dim(Ncon)[[1]]+dim(Ncon)[[2]]+1):dim(mat)[[2]]] <- as.matrix(Ncon)

# convert adjacency matrix into graph object
g2 <- graph.adjacency(mat,mode="directed",weighted=T)

# extract weights
#wts <- g2[[9]][[4]]$weight
wts <- E(g2)$weight

# create layout matrix
locs <- cbind(c(phage.time,bacteria.time),c(rep(0,length(phage.time)),rep(-1,length(bacteria.time)/2),rep(1,length(bacteria.time)/2)))

#plot options
phage.color = "black"; bac.color = "white"; bac.color2 = "grey"
#coevo.shape = "circle"; control.shape = "square"


# create plot
plot(g2,
     #vertex.size = sq.dim,
     #vertex.label.color = "black",
     vertex.label = NA,
     vertex.shape = c(rep("square",length(phage.time)),
                      rep("circle",length(bacteria.time)/2),
                      rep("circle",length(bacteria.time)/2)),
     vertex.color = c(rep(phage.color,length(phage.time)),
                      rep(bac.color,length(bacteria.time)/2),
                      rep(bac.color2,length(bacteria.time)/2)),
     edge.color = "grey25",edge.width = 8*wts,
     edge.arrow.mode = 0,layout = locs,
     xlab = list("Time (d)", cex = 1.5, font = 2),asp = 0, ylim = c(-1.2,1))

# draw scaled axes
xs <- c(0,40,80, 120, 160)
scaled.xs <- 2*(xs - min(locs[,1]))/(max(locs[,1]) - min(locs[,1])) - 1

axis(1,at = scaled.xs,labels = xs, cex = 1.15)
axis(2, at = c(-1,0,1),labels = c("+Ph","phage","-Ph"),cex.axis = 1.15,font = 1, las = 1)

box(lwd = 3)


### PL treatment
#Pcon
#Ptrt

#Option 1
#Merge data from each matrix
#data.in <- rbind(Pcon, Ptrt)
#data.in <- as.matrix(data.in)

#generate a square adjacency matrix, creating a place to store the data
#mat <- matrix(0,nrow=sum(dim(data.in)),ncol=sum(dim(data.in)))
  
#nrows <- dim(data.in)[[1]]
#big.nrows <- dim(mat)[[2]]
  
# create the phage-bacteria edges between time points
#mat[1:nrows,I(nrows+1):big.nrows] <- data.in

# convert adjacency matrix into graph object
#g <- graph.adjacency(mat,mode = "directed",weighted = T)

# extract weights
#wts <- g2[[9]][[4]]$weight
#wts <- E(g)$weight

#Option 2
# construct and populate adjacency matrix
size <- 1*dim(Pcon)[[1]] + 2*dim(Ptrt)[[2]]
mat <- matrix(0, nrow = size, ncol = size)
#mat

# for multiple bacteria, single phage 
mat[1:dim(Pcon)[[1]],I(dim(Pcon)[[1]]+1):I(dim(Pcon)[[1]] + dim(Pcon)[[2]])] <- as.matrix(Ptrt)
mat[1:dim(Pcon)[[1]],I(dim(Pcon)[[1]]+dim(Pcon)[[2]]+1):dim(mat)[[2]]] <- as.matrix(Pcon)

# convert adjacency matrix into graph object
g2 <- graph.adjacency(mat,mode="directed",weighted=T)

# extract weights
#wts <- g2[[9]][[4]]$weight
wts <- E(g2)$weight

# create layout matrix
locs <- cbind(c(phage.time,bacteria.time),c(rep(0,length(phage.time)),rep(-1,length(bacteria.time)/2),rep(1,length(bacteria.time)/2)))

#plot options
phage.color = "black"; bac.color = "white"; bac.color2 = "grey"
coevo.shape = "circle"; control.shape = "square"

# create plot
plot(g2,
     #vertex.size = sq.dim,
     #vertex.label.color = "black",
     vertex.label = NA,
     vertex.shape = c(rep("square",length(phage.time)),
                      rep("circle",length(bacteria.time)/2),
                      rep("circle",length(bacteria.time)/2)),
     vertex.color = c(rep(phage.color,length(phage.time)),
                      rep(bac.color,length(bacteria.time)/2),
                      rep(bac.color2,length(bacteria.time)/2)),
     edge.color = "grey25",edge.width = 8*wts,
     edge.arrow.mode = 0,layout = locs,
     xlab = list("Time (d)", cex = 1.5, font = 2),
     asp = 0,ylim = c(-1.2,1))

# draw scaled axes
xs <- c(0,40,80, 120, 160)
scaled.xs <- 2*(xs - min(locs[,1]))/(max(locs[,1]) - min(locs[,1])) - 1

axis(1,at = scaled.xs,labels = xs, cex = 1.15)
axis(2, at = c(-1,0,1),labels = c("+Ph","phage","-Ph"),cex.axis = 1.15,font = 1, las = 1)

box(lwd = 3)

dev.off()

```

## Evolutionary time shifts

```{r fig.height = 5, fig.width = 5}


## This calculation is modified from Gaba et al
bor.cstat$ptime.adj <- bor.cstat$phage.time
bor.cstat$btime.adj <- bor.cstat$bac.time

bor.cstat$btime.adj[bor.cstat$btime.adj == -6] <- 1
bor.cstat$btime.adj[bor.cstat$btime.adj == 9] <- 2
bor.cstat$btime.adj[bor.cstat$btime.adj == 23] <- 3
bor.cstat$btime.adj[bor.cstat$btime.adj == 72] <- 4
bor.cstat$btime.adj[bor.cstat$btime.adj == 100] <- 5
bor.cstat$btime.adj[bor.cstat$btime.adj == 129] <- 6
bor.cstat$btime.adj[bor.cstat$btime.adj == 166] <- 7
bor.cstat$ptime.adj[bor.cstat$ptime.adj == 9] <- 2
bor.cstat$ptime.adj[bor.cstat$ptime.adj == 23] <- 3
bor.cstat$ptime.adj[bor.cstat$ptime.adj == 72] <- 4
bor.cstat$ptime.adj[bor.cstat$ptime.adj == 129] <- 6
bor.cstat$ptime.adj[bor.cstat$ptime.adj == 166] <- 7

bor.cstat$time.shift <- bor.cstat$phage.time-bor.cstat$bac.time
bor.cstat$bac.time.fac <- as.factor(bor.cstat$bac.time)
bor.cstat$p.time.shift <- bor.cstat$btime.adj-bor.cstat$ptime.adj
bor.cstat$phage.time.fac <- as.factor(bor.cstat$phage.time)


facet_names <- list(
  'N'="N-limited",
  'P'="P-limited",
  'C'="Non-phage amended control",
  'T'="Phage amended"
)

facet_labeller <- function(variable,value){
  return(facet_names[value])
}


# Time shift by time point
p <- ggplot(bor.cstat, aes(x = time.shift, y = inf.prob, col = bac.time.fac, label = bac.time.fac)) +
  #geom_point()+
#  geom_point(bor.cstat[bor.cstat$time.shift == 0,], aes(x = time.shift, y = inf.prob),
#             col = "black",size = 1.25) +
  #annotate("text", x = label = bac.time.fac) +
  geom_vline(xintercept = 0, lty = 2, col = "grey50")+
  stat_summary(fun.y = mean,geom="line")+
  facet_grid(trt~lim, labeller=facet_labeller)+
  xlab("Time Shift (days) relative to host") +
  ylab("Interaction Strength") +
  theme_bw()+
  theme(legend.position='none',
        axis.text.x  = element_text(size=10),
        axis.text.y  = element_text(size=10),
        axis.title.x = element_text(colour="black", size=12,
                                    margin = margin(20,20,10,10)),
        axis.title.y = element_text(colour="black", size=12,
                                    margin = margin(20,20,10,10)),
        panel.grid = element_blank(),
        strip.background = element_rect(fill="white"))
print(p)


#ggsave("./supporting-files/time_shift.tif", p, device = "tiff", 
#       scale = 1, width = 6, height = 4, units = "in", dpi = 300)


p1 <- ggplot(bor.cstat, aes(x = time.shift, y = inf.prob)) +
  geom_point(aes(col = bac.time.fac)) + 
  #stat_summary(fun.y = mean, geom="line")+
  facet_grid(trt~lim)+
  stat_smooth(geom = "smooth", method = "loess")

p2 <- ggplot(bor.cstat, aes(x = p.time.shift, y = inf.prob, col = phage.time.fac)) +
  geom_point() + stat_summary(fun.y = mean, geom="line")+
  facet_grid(trt~lim)

p3 <- ggplot(bor.cstat, aes(x = p.time.shift, y = inf.prob)) +
  geom_point(aes(col = phage.time.fac)) + stat_summary(fun.y = mean, geom="line")+
  facet_grid(trt~lim)


#Run RMANOVA with different models
model.ar <- lme(inf.prob ~ trt * lim * time.shift, random = ~1 | BcID,
                 correlation = corAR1(form = ~1 ),
                 data = bor.cstat)

model.arma <- lme(inf.prob ~ trt* lim * time.shift,random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 1),
                 data = bor.cstat)

model.arma1 <- lme(inf.prob ~ trt* lim * time.shift, random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 2),
                 data = bor.cstat)

#Select which model fits the data better
#anova(model.ar,model.arma, model.arma1)
pander(summary(model.arma1))

#Run RMANOVA for just controls
model.arma1.controls <- lme(inf.prob ~ lim * time.shift, random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 2),
                 data = bor.cstat[bor.cstat$trt == "C",])
summary(model.arma1.controls)
anova(model.arma1.controls)

model.arma1.treatment <- lme(inf.prob ~ lim * time.shift, random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 2),
                 data = bor.cstat[bor.cstat$trt == "T",])
summary(model.arma1.treatment)
anova(model.arma1.treatment)
```

\newpage
```{r fig.height = 5, fig.width = 5}

b.type.func <- function(x,y){
  if(y < x){
    type <- "P"
    } else if(y == x){
      type <- "C"
      } else {
        type <- "F"
      }
  return(type)
  }

bor.cstat$b.type <- mapply(b.type.func,bor.cstat$bac.time,bor.cstat$phage.time)
bor.cstat$p.type <- mapply(b.type.func,bor.cstat$phage.time, bor.cstat$bac.time)

# cyano time
p4 <- ggplot(bor.cstat, aes(x = as.factor(bac.time), inf.prob)) +
  geom_point(aes(shape = as.factor(phage.time), color = BcID)) +
  stat_summary(fun.data = "mean_cl_boot", col = "black", size = 0.75) + 
  facet_grid(trt~lim) +
    theme_bw()+
  theme(legend.position='none',
        axis.text.x  = element_text(size=10),
        axis.text.y  = element_text(size=10),
        axis.title.x = element_text(colour="black", size=12,
                                    margin = margin(20,20,10,10)),
        axis.title.y = element_text(colour="black", size=12,
                                    margin = margin(20,20,10,10)),
        panel.grid = element_blank()) +
  xlab("Synechococcus Isolation Time (d)") +
  ylab("Interaction Strength")

print(p4)

ggsave("./supporting-files/cyano_time.tif", p4, device = "tiff", 
      scale = 1, width = 6, height = 4, units = "in", dpi = 300)
```

```{r}
fit <- aov(inf.prob ~ bac.time.fac * lim * trt, data = bor.cstat)
summary(fit)
TukeyHSD(fit)

fit <- aov(inf.prob ~ bac.time.fac * lim, data = bor.cstat[bor.cstat$trt == "C",])
summary(fit)
TukeyHSD(fit)


model.ar <- lme(inf.prob ~ trt * lim * bac.time, random = ~1 | BcID,
                 correlation = corAR1(form = ~1 ),
                 data = bor.cstat)

model.arma <- lme(inf.prob ~ trt* lim *  bac.time,random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 1),
                 data = bor.cstat)

model.arma1 <- lme(inf.prob ~ trt* lim * bac.time, random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 2),
                 data = bor.cstat)

#Select which model fits the data better
#anova(model.ar,model.arma, model.arma1)
pander(summary(model.arma1))

model.arma1T <- lme(inf.prob ~ lim * bac.time, random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 2),
                 data = bor.cstat[bor.cstat$trt == "T",])
pander(summary(model.arma1T))
stats = anova.lme(model.arma1P,type = "sequential", adjustSigma = F)

model.arma1C <- lme(inf.prob ~ lim * bac.time, random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 2),
                 data = bor.cstat[bor.cstat$trt == "C",])
pander(summary(model.arma1C))

model.arma1N <- lme(inf.prob ~ trt * bac.time, random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 2),
                 data = bor.cstat[bor.cstat$lim == "N",])
pander(summary(model.arma1N))

model.arma1P <- lme(inf.prob ~ trt * bac.time, random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 2),
                 data = bor.cstat[bor.cstat$lim == "P",])
pander(summary(model.arma1P))


```

\newpage

```{r fig.height = 5, fig.width = 5}

#phage time

p5 <- ggplot(bor.cstat, aes(x = as.factor(phage.time), inf.prob)) +
  geom_point(aes(shape = as.factor(phage.time), color = BcID)) +
  stat_summary(fun.data = "mean_cl_boot", col = "black", size = 0.75) + 
  facet_grid(trt~lim) +
    theme_bw()+
  theme(legend.position='none',
        axis.text.x  = element_text(size=10),
        axis.text.y  = element_text(size=10),
        axis.title.x = element_text(colour="black", size=12,
                                    margin = margin(20,20,10,10)),
        axis.title.y = element_text(colour="black", size=12,
                                    margin = margin(20,20,10,10)),
        panel.grid = element_blank()) +
  xlab("Phage Isolation Time (d)") +
  ylab("Interaction Strength")

ggsave("./supporting-files/phage_time.tif", p5, device = "tiff", 
      scale = 1, width = 6, height = 4, units = "in", dpi = 300)

aggregate(inf.prob ~ trt + lim + phage.time, data = bor.cstat, mean)

model.arma1 <- lme(inf.prob ~ trt* lim * phage.time, random = ~1 | BcID,
                 correlation = corARMA(form = ~1, p = 1, q = 2),
                 data = bor.cstat)
summary(model.arma1)

p <- ggplot(bor.cstat, aes(x = b.type, inf.prob)) +
  geom_boxplot(aes(fill = bac.time.fac)) +
  facet_grid(trt~lim)

p$data$b.type <- factor(p$data$b.type, levels = c("P","C","F"))
p$data$bac.time <- factor(p$data$bac.time, levels = c("-6","9","23","72","100","129","166"))
#print(p)


p <- ggplot(bor.cstat, aes(x = bac.time.fac, inf.prob)) +
  geom_boxplot(aes(fill = b.type), size = 1) +
  facet_grid(trt~lim)

p$data$b.type <- factor(p$data$b.type, levels = c("P","C","F"))
p$data$bac.time <- factor(p$data$bac.time, levels = c("-6","9","23","72","100","129","166"))
#print(p)

# Bact coevol model
p <- ggplot(bor.cstat, aes(x = b.type, inf.prob)) +
  geom_boxplot() +
  facet_grid(trt~lim) +
  xlab("Bacteria Time") +
  ylab("Interaction Strength")
p$data$b.type <- factor(p$data$b.type, levels = c("P","C","F"))
#print(p)

# Phage coevol model
p1 <- ggplot(bor.cstat, aes(x = p.type, inf.prob)) +
  geom_boxplot() +
  facet_grid(trt~lim) +
  xlab("Phage Time") +
  ylab("Interaction Strength")
p1$data$p.type <- factor(p1$data$p.type, levels = c("P","C","F")) 
#print(p1)

grid.arrange(p,p1, ncol = 2)

aggregate(inf.prob ~ trt + lim * bac.time, data = bor.cstat, sd)

```
